The Splashback Radius of Planck-SZ clusters

### 2019-01-24 Paper_v2/RSP_paper_new.tex
Some edits

\textbf{Trying to avoid the problems which complicate the unbiased detction
of the splashback signal in opical samples some recent studies searched
for the signature of the splashback effect using non-optical cluster samples.}
\citet{umetsu2017lensing} as well as \citet{contigiani2018weak}
used Xray selected clusters to look for the splashback radius
using the weak lensing signal, however stacking issues and the low
signal-to-noise ratio remains a significant hurdle for both of them.
Also \citet{chang2017splashback} found evidence for the splashback feature
in the weak lensing signal, but their analysis was again done using
optically selected clusters in the Dark energy survey suffering from 
the before mentioned issus again.

\textbf{In this work, we explore yet another non-optical cluster sample. 
Namely, SZ selected clusters.} While the SZ
selected clusters can also be susceptible to systematic selection
effects, the scales on which the SZ signal is measured and the
cluster selection is performed is much smaller than the expected
location of the splashback radius (typically $R_{\rm 500c}$). We use
this sample to explore the evidence for the splashback radius in
observations.  Due to the low signal-to-noise ratio of the weak
lensing signal, we perform our analysis using cross-correlation of
galaxies with clusters.

\textbf{Dynamcial friction is expected to decrease the observed splashback radius in 
massive clusters. It causes a
slow-down of the massive subhalos, hosting many of the observed galaxies, 
which in return decreases the apocentric
radius of their orbits and therefore the measured splashback radius of the host cluster. 
Since the rate of dynamical friction scales
with the mass of the moving object the effect is stronger for more
massive subhalos. Therefore, one expects the decrease in the splashback
radius to scale with the amount of massive subhalos in the host cluster and
with its mass \citep{adhikari2016observing}. This bias can be limited by choosing a deep enough
galaxy sample such that a large enough amount of low mass subhalos can be
observed.} The SDSS sample used by
\citet{more2016detection} was already deep enough such that biases in the
location of the splashback radius due to dynamical friction effects
were expected to be small. 
Nevertheless, we use galaxy samples, which
are even fainter by $0.5-1$ magnitudes compared to those used by
\citet{more2016detection}.

@@

Therefore the model is described by eight model parameters.
We infer the posterior distributions of those parameters by fitting the model for the two-dimensional correlation
function to the measured two-dimensional correlation signal. 
We use the affine invariant Markov Chain Monte Carlo sampler of \citet{goodman2010ensemble} as
implemented in the parallel python package {\it emcee} by
\citet{foreman2013emcee}. 
\textbf{We keep the parameters $\rho_{\mathrm{s}}$, $\rho_0$ and $s_{\mathrm{e}}$ 
as free parameters during the fitting procedure. Similarly to \citet{more2016detection} 
we put flat priors on $\log_{10}(r_{\mathrm{s}})$ and $\log_{10}(r_{\mathrm{t}})$ but double 
their ranges to $[0.1,5.0]$ compared to their work. In contrast, 
we use a normal prior with a central value of $\log_{10}(0.2)$ and a scale of 1.2 
for $\log_{10}(\alpha)$ as deduced from mass estimates \citep{gao2008redshift}. 
For the parameters $\log_{10}(\beta)$ and $\log_{10}(\gamma)$ we also use normal 
priors with central values $\log_{10}(6.0)$ and $\log_{10}(4.0)$, respectively 
and scale 0.4 as recommended by \citet{diemer2014dependence}.}

\textbf{It has been a common concern that the profile in Equation~\ref{eq:model} might introduce
a splashback feature simply by its functional from and without its presence in the data.
However, \citet{baxter2017halo} showed that this is not the case by using a Bayesian odds 
ratio approach to asses that the data is described significantly better by the general 
profile in Equation~\ref{eq:model} (with the splashback feature)
rather than by the same profile with fixed $f_{trans} = 1$ (without splashback feature).
The same conclusion was made by \citet{umetsu2017lensing} who conducted a similar study but
using weak lensing data.}
%\begin{table*}
%    \centering
%    \caption{Summary of the priors used in the MCMC sampling
%procedure. The MCMC chains are constrained using flat priors or
%normal priors on some of the fitting parameters. This table lists
%the ranges of the flat priors and the central positions as well as
%the scales of the normal priors ( format: center | scale ),
%respectively.}
%    \label{tab:priors}
%    \begin{tabular}{ccccccccc}
%    \hline 
%    & $\log_{10}(\rho_{\mathrm{s}})$ & $\log_{10}(\alpha)$ & $\log_{10}%(r_{\mathrm{s}})$ & $\rho_0$ & $s_{\mathrm{e}}$ & $\log_{10}(r_{\mathrm{t}})$ & $\log_{10}(\beta)$ & $\log_{10}(\gamma)$ \\ 
%    \hline 
%    Prior Type & None & Normal & Flat & None & None & Flat & Normal & Normal\\ 
%    \hline 
%    Prior Range & - & $\log_{10}(0.2)| 1.2$ & [0.1,5.0] & - & - & [0.1,5.0] & $%\log_{10}(6.0) | 0.4$ & $\log_{10}(4.0) | 0.4$ \\
%    \hline
%    \end{tabular} 
%\end{table*}

\textbf{While there have been other suggestions for defining the location of the
splashback radius (see \citet{adhikari2014splashback, shi2016outer, 
mansfield2017splashback, diemer2017splashback}) we adopt the 
definition given by \citet{more2015splashback}, who define
the location of the splashback radius for halos on galaxy cluster scales to 
be consistent with the location of the steepest logarithmic slope of the density
profile, due to its easy accessibility in data.}

@@

\section{Results}
\label{sec:Results}
The measurements of the two-dimensional cluster-galaxy
cross-correlations are shown as black points with error bars in the
first row of Figure~\ref{fig:all_graphs}. The different columns
correspond to the three different absolute magnitude limits that we
have used to select all the galaxies when calculating the
cross-correlations. The cross-correlation signal is clearly detected
in all three measurements. \textbf{The signal to noise ratios of the 
three different estimates of the two-dimensional cross-correlation 
signals amount to 42.4, 43.3 and 30.9 for the galaxy samples PS 21, 
PS 21.5 and PS 22, respectively. For comparison: 
The SNR achieved by \citet{more2016detection} is 263. It is much higher
due to size of the cluster sample used in their study.}

@@

\textbf{We show the two-dimensional posterior
distributions for each pair of parameters for the PS 21.5 sample in 
Figure~\ref{fig:corner_21.5}. Since the posterior distributions for 
the other datasetsare very similar we refrain from showing them as well. 
It is observed that there are some strong degeneracies present between 
parameters describing the same part of the model in Equation~\ref{eq:model} 
(namely $\rho_{\mathrm{in}}$, $f_{\mathrm{trans}}$ and $\rho_{\mathrm{out}}$).
This is not unexpected since certain combinations of parameter values can 
very well result in the same functional shape of the different parts of 
the model. This degeneracy also leads to some of the parameters not being 
constrained tightly (especially $\rho_s$ and $r_s$ due to their strong degeneracy 
with $\alpha$). On the other hand, it is important to note that there is 
no degeneracy present between parameters describing different parts of the 
model, which means that the overall shape of the profile is well constrained. 
Note also that, since the location of the splashback radius only depends on 
the overall shape of the profile the uncertainties of the individual parameters 
within the independent parts of the profile do not translate
into an uncertainty of the location of the splashback feature directly.}

@@

\textbf{This result is in contradiction to the findings of \citet{baxter2017halo}, who find no splashback feature in the blue galaxy population. At the current state it is not clear how those different results emerge. It might be that, in contrast to our SZ selected cluster sample, there is simply no splashback feature present in the galaxy cluster sample used by \citet{baxter2017halo}, which was constructed using the REDMAPPER algorithm on SDSS data. Further, the separation of the galaxies by color used by \citet{baxter2017halo} is fundamentaly different from our approach. They use K-corrected magnitudes from the SDSS dataset, whereas we avoid using K-corrections but instead use a redshift dependent color cut.}

>>>>

\textbf{To avoid optical cluster selection effects, there have been some
previous efforts to use X-ray selected cluster samples to look for the
splashback radius using the weak lensing signal\{umetsu2017lensing,
contigiani2018weak}}.  However stacking issues and the low signal-to-noise
ratio remains a significant hurdle for both of them. \textbf{The use of weak
lensing in these studies could get around the biased nature of the galaxy
distribution by directly probing the dark matter distribution. Along those
lines,} \citet{chang2017splashback} found evidence for the splashback feature
in the weak lensing signal, but their analysis was again done using optically
selected clusters in the Dark energy survey that suffer from the
afore-mentioned issues.

\textbf{In this work, we explore clusters selected using the SZ effect.} While
the SZ selected clusters can also be susceptible to systematic selection
effects, the scales on which the SZ signal is measured and the cluster
selection is performed is much smaller than the expected location of the
splashback radius (typically $R_{\rm 500c}$). We use this sample to explore the
evidence for the splashback radius in observations.  Due to the low
signal-to-noise ratio of the weak lensing signal, we perform our analysis using
cross-correlation of galaxies with clusters.

\textbf{Dynamical friction is expected to slow down massive subhalos, hosting
galaxies that are orbiting the galaxy cluster XXXX{Cite paper by Chandrasekhar
here}XXXX. If a majority of the observed galaxies are hosted by these massive
subhalos, the slow down due to dynamical friction would reduce the apocenters
of their orbits, in turn, biasing the measured splashback radius of the host
cluster XXX{Cite More2016 and papers on Benedikt's SPARTA}XXX. As dynamical
friction timescales are inversely proportional to the mass of the subhalos,
this effect is stronger for more massive subhalos. Therefore, one expects the
splashback radius to decrease for more massive subhalos
\citep{adhikari2016observing}.  This bias can be limited by choosing a faint
galaxy sample such that majority of the galaxies reside in low mass subhalos,
unaffected by dynamical friction.} The SDSS sample used by
\citet{more2016detection} was already deep enough such that biases in the
location of the splashback radius due to dynamical friction effects were
expected to be small. Nevertheless, we use galaxy samples, which are fainter by
$0.5-1$ magnitudes compared to those used by \citet{more2016detection}.

@@

Therefore the model is described by eight model parameters.
We infer the posterior distributions of those parameters by fitting the model for the two-dimensional correlation
function to the measured two-dimensional correlation signal. 
We use the affine invariant Markov Chain Monte Carlo sampler of \citet{goodman2010ensemble} as
implemented in the parallel python package {\it emcee} by
\citet{foreman2013emcee}. 
\textbf{We let the parameters $\rho_{\mathrm{s}}$, $\rho_0$ and
$s_{\mathrm{e}}$ vary freely during our fitting procedure. Similar to
\citet{more2016detection} we put flat priors on $\log_{10}(r_{\mathrm{s}})$ and
$\log_{10}(r_{\mathrm{t}})$ but double their ranges to $[0.1,5.0]$ compared to
their work. In contrast to \citet{more2016detection}, we use a normal prior
with a central value of $\log_{10}(0.2)$ and a scale of 1.2 for
$\log_{10}(\alpha)$ as deduced from mass estimates \citep{gao2008redshift}.
For the parameters $\log_{10}(\beta)$ and $\log_{10}(\gamma)$ we also use
normal priors with central values $\log_{10}(6.0)$ and $\log_{10}(4.0)$,
respectively and scale 0.4 as recommended by \citet{diemer2014dependence}.}

\textbf{It has been a common concern that the profile in
Equation~\ref{eq:model} introduces a point of minimum slope in the logarithmic
derivative profile simply due to the functional form. However,
\citet{baxter2017halo} used a Bayesian evidence approach to assess that the
data is described significantly better by the general profile in
Equation~\ref{eq:model} (with the splashback feature) rather than by the same
profile with fixed $f_{trans} = 1$ (without splashback feature).  The same
conclusion was made by \citet{umetsu2017lensing} who conducted a similar study
but using weak lensing data.}
%\begin{table*}
%    \centering
%    \caption{Summary of the priors used in the MCMC sampling
%procedure. The MCMC chains are constrained using flat priors or
%normal priors on some of the fitting parameters. This table lists
%the ranges of the flat priors and the central positions as well as
%the scales of the normal priors ( format: center | scale ),
%respectively.}
%    \label{tab:priors}
%    \begin{tabular}{ccccccccc}
%    \hline 
%    & $\log_{10}(\rho_{\mathrm{s}})$ & $\log_{10}(\alpha)$ & $\log_{10}%(r_{\mathrm{s}})$ & $\rho_0$ & $s_{\mathrm{e}}$ & $\log_{10}(r_{\mathrm{t}})$ & $\log_{10}(\beta)$ & $\log_{10}(\gamma)$ \\ 
%    \hline 
%    Prior Type & None & Normal & Flat & None & None & Flat & Normal & Normal\\ 
%    \hline 
%    Prior Range & - & $\log_{10}(0.2)| 1.2$ & [0.1,5.0] & - & - & [0.1,5.0] & $%\log_{10}(6.0) | 0.4$ & $\log_{10}(4.0) | 0.4$ \\
%    \hline
%    \end{tabular} 
%\end{table*}

\textbf{The location of the splashback radius can be identified in various
ways, e.g., as the radius of the outermost shell in theoretical models of
spherical symmetric collapse \citep{adhikari2014splashback, shi2016outer}, or
using non-spherical shells of density jumps \citet{ mansfield2017splashback},
or the average location of the apocenter of particles falling in the cluster
potential \citet{diemer2017splashback}. Although these definitions have subtle
differences, they all follow the general trend of a decreasing splashback
radius with increasing rate of accretion XXX{Cite Benedikt and Phils paper for
these comparisons}. In this paper, we adopt the convention of
\citet{more2015splashback}, who define the location of the splashback radius
for halos on galaxy cluster scales to be the location of the steepest
logarithmic slope of the radially averaged density profile, due to its ease of
accessibility in data.}

@@

\section{Results}
\label{sec:Results}
The measurements of the two-dimensional cluster-galaxy
cross-correlations are shown as black points with error bars in the
first row of Figure~\ref{fig:all_graphs}. The different columns
correspond to the three different absolute magnitude limits that we
have used to select all the galaxies when calculating the
cross-correlations. The cross-correlation signal is clearly detected
in all three measurements. \textbf{The signal to noise ratios of the 
three different estimates of the two-dimensional cross-correlation signals
amount to 42.4, 43.3 and 30.9 for the galaxy samples PS 21, PS 21.5 and PS 22,
respectively. For comparison, the detection SNR in \citet{more2016detection}
was 263, a much higher number due to the size of the cluster sample used in
their study.}

@@

\textbf{We show the two-dimensional posterior distributions for each pair of
parameters for the PS 21.5 sample in Figure~\ref{fig:corner_21.5}. The
posterior distributions for the other datasets show similar degeneracies and
are not included in the paper for brevity.
It is observed that there are some strong degeneracies present between 
parameters describing the same part of the model in Equation~\ref{eq:model} 
(namely $\rho_{\mathrm{in}}$, $f_{\mathrm{trans}}$ and $\rho_{\mathrm{out}}$).
This is expected since certain combinations of parameter values can 
very well result in the same functional shape of the different parts of 
the model. This degeneracy also leads to some of the parameters not being 
constrained tightly (especially $\rho_s$ and $r_s$ due to their strong degeneracy 
with $\alpha$). On the other hand, it is important to note that there is
little-to-no degeneracy present between parameters describing different parts
of the model, which means that the overall shape of the profile is well
constrained.  Note also that, since the location of the splashback radius only
depends on the overall shape of the profile, the uncertainties of the
individual parameters within the independent parts of the profile do not
translate into a huge uncertainty on the location of the splashback feature.}

@@

\textbf{This result seem to be at odds with the findings presented by
\citet{baxter2017halo}, who find no splashback feature in the blue galaxy
population. Our investigations suggest that the splashback feature gets less
prominent as we proceed to look at galaxies which are progressively bluer in
color. Given that \citet{baxter2017halo}, consider the last quartile of the
blue population, it is imaginable that they see no evidence for the splashback
feature in their dataset. The approach used by \citet{baxter2017halo} used to
separate galaxies into red or blue is also fundamentally different from our
approach. While \citet{baxter2017halo} use K-corrected magnitudes from SDSS
relying on the photometric redshifts of galaxies (which could have different
uncertainties for the red and the blue population), we rely on a color cut
which depends upon the redshift of the cluster under consideration.}
